---
title: "Prediction Report: Exercise Classification"
author: "Milda"
date: "`r Sys.Date()`"
output: html_document
---

# Introduction
In this project, we predict the manner in which exercises were performed using a dataset collected from wearable sensors. The target variable is **classe**, which categorizes the type of exercise performed. This report details the data preprocessing, model training, validation, and final predictions.

# Data Loading & Preprocessing
The dataset is loaded, and columns with a high proportion of missing values are removed. Unnecessary metadata columns are also eliminated.

```{r, message=FALSE, warning=FALSE}
library(caret)
library(randomForest)
library(ggplot2)
library(dplyr)
library(readr)

train_path <- "../data/pml-training.csv"
test_path <- "../data/pml-testing.csv"

training <- read_csv(train_path)
testing <- read_csv(test_path)

# Remove columns with excessive missing values
na_threshold <- 0.9
na_counts <- colSums(is.na(training)) / nrow(training)
training <- training[, na_counts < na_threshold]

# Ensure testing dataset has the same features as training (excluding 'classe')
common_features <- setdiff(names(training), "classe")
testing <- testing[, common_features, drop = FALSE]

# Remove irrelevant metadata columns
training <- training[, -c(1:7)]
training$classe <- as.factor(training$classe)
```

# Model Training
We use the **Random Forest** algorithm to train the model with 70% of the training data. The model predicts the target variable `classe`.

```{r}
set.seed(123)
trainIndex <- createDataPartition(training$classe, p = 0.7, list = FALSE)
trainData <- training[trainIndex, ]
validData <- training[-trainIndex, ]

rf_model <- randomForest(classe ~ ., data = trainData, ntree = 100)
```

# Model Validation
The trained model is validated against the remaining 30% of the data, and accuracy is measured using a confusion matrix.

```{r}
rf_predictions <- predict(rf_model, validData)
conf_matrix <- confusionMatrix(rf_predictions, validData$classe)
conf_matrix
```

# Feature Importance
Random Forest provides feature importance scores, which help understand the significance of each variable in the classification.

```{r}
importance <- varImp(rf_model)
plot(importance, main = "Feature Importance")
```

# Predictions on Test Data
The trained model is applied to the test dataset, and the predictions are saved in individual text files.

```{r}
final_predictions <- predict(rf_model, testing)

prediction_files <- function(predictions) {
  for (i in 1:length(predictions)) {
    filename <- paste0("../results/problem_id_", i, ".txt")
    write.table(predictions[i], file = filename, quote = FALSE, row.names = FALSE, col.names = FALSE)
  }
}
prediction_files(final_predictions)
```

# Conclusion
- The **Random Forest model** was successfully trained and validated.
- The model achieved high accuracy on the validation dataset.
- Predictions for the test dataset were saved in the required format.
- Further optimization and feature selection could improve model performance.
